### Analyze Zika infected Glioblastoma data:

---
  output:
  html_document: default
pdf_document: default
---
  ### Finding gene modules that make cells susceptible to Zika infection
  
  ---
  title: "Finding gene modules that make cells susceptible to Zika infection"
output:
  html_document: default
github_document: default
fig_width: 12
fig_height: 4 
---
  
```{r include = FALSE}
### Load data and packages
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

Load the required R libraries:
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
library(Seurat)
library(Matrix)
library(rhdf5)
library(dplyr)
require(data.table)
```

Prior to this analysis we processed the raw sequencing data with cellranger and ran an algorithm called CellBender on it to remove ambient RNA.

Here I 1.) load the data 2.) change the gene identifiers from ensemble ids to symbols 3.) construct a metadata matrix (with donor information, pcw etc.)  4.) put the data and metadata into a Seurat object. 5.) load mitochondrial genes from a database for later use 6.) Remove doublets, by loading the doublet scores I got from an algorithm (called scrublet) I ran on the data before.
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
setwd('/home/jovyan/AA_DOW/')
# sample_info = data.frame(matrix('', 8,4)) # (info from sample tracker)
# colnames(sample_info) = c('SangerSampleID', 'Donor', 'Disease', 'PcW')
# sample_info$SangerSampleID = c('DOW8726784', 'DOW8726783', 'DOW8726787', 'DOW8726788', 'DOW8726785', 'DOW8726786', 'DOW8726789', 'DOW8726790')
# sample_info$Donor = c('Hob1', 'Hob1', 'Hob12', 'Hob12', 'Hob15', 'Hob15', 'Hob16', 'Hob16')
# sample_info$Disease = c('TS21', 'TS21', 'TS21', 'TS21', 'Control', 'Control', 'Control', 'Control')
# sample_info$PcW = c('13pcw', '13pcw', '14pcw', '14pcw', '13pcw', '13pcw', '13pcw', '13pcw')
# scrublet_cutoff = c(0.33, 0.25, 0.4, 0.4, 0.4, 0.45, 0.4, 0.3) # From inspection of the scrublet plots (not included). This is quite a stringent doublet cutoff.
# i = 1
# print(i)
# sampleID = sample_info$SangerSampleID[i]
# scrublet_score = read.delim(paste('/home/jovyan/AA_DOW/scrublet/', sampleID ,'_doublets_scores.txt', sep = ''))
# dow_counts = as.matrix(Read10X_h5(filename = paste('../data/AA_DOW/AA_DOW/cellranger302_count_33315_AA_', sampleID, '_GRCh38-3_0_0-premrna/output_filtered.h5', sep = ''), use.names = TRUE))
# dow_counts = dow_counts[,scrublet_score < scrublet_cutoff[i]]
# dow_metadata = data.frame(matrix('',dim(dow_counts)[2],4))
# colnames(dow_metadata) = c('SangerSampleID', 'Donor', 'Disease', 'PcW')
# dow_metadata$SangerSampleID = sample_info$SangerSampleID[i]
# dow_metadata$Donor = sample_info$Donor[i]
# dow_metadata$Disease = sample_info$Disease[i]
# dow_metadata$PcW = sample_info$PcW[i]
# geneNames = read.delim('/home/jovyan/data/AA_DOW/AA_DOW/cellranger302_count_33315_AA_DOW8726783_GRCh38-3_0_0-premrna/raw_feature_bc_matrix/features.tsv')
# for (i in 2:8){
#   print(i)
#   sampleID = sample_info$SangerSampleID[i]
#   scrublet_score = read.delim(paste('/home/jovyan/AA_DOW/scrublet/', sampleID ,'_doublets_scores.txt', sep = ''))
#   dow_counts_subset = as.matrix(Read10X_h5(filename = paste('../data/AA_DOW/AA_DOW/cellranger302_count_33315_AA_', sampleID, '_GRCh38-3_0_0-premrna/output_filtered.h5', sep = ''), use.names = TRUE))
#   dow_counts_subset = dow_counts_subset[,scrublet_score < scrublet_cutoff[i]]
#   dow_metadata_subset = data.frame(matrix('',dim(dow_counts_subset)[2],4))
#   colnames(dow_metadata_subset) = c('SangerSampleID', 'Donor', 'Disease', 'PcW')
#   dow_metadata_subset$SangerSampleID = sample_info$SangerSampleID[i]
#   dow_metadata_subset$Donor = sample_info$Donor[i]
#   dow_metadata_subset$Disease = sample_info$Disease[i]
#   dow_metadata_subset$PcW = sample_info$PcW[i]
#   dow_counts = cbind(dow_counts, dow_counts_subset)
#   dow_metadata = rbind(dow_metadata, dow_metadata_subset)
# }
# mitogenes = geneNames[,2][substring(geneNames[,2],1,3) == 'MT-']
# percent.mt = colSums(dow_counts[rownames(dow_counts) %in% mitogenes,])/colSums(dow_counts)
# Downs <- CreateSeuratObject(dow_counts, project = 'AA_DOW', min.cells = 0, min.features = 0)
# Downs$Barcode = colnames(dow_counts)
# Downs$SangerSampleID = dow_metadata$SangerSampleID
# Downs$Donor = dow_metadata$Donor
# Downs$Disease = dow_metadata$Disease
# Downs$percent.mt = unname(percent.mt)
# saveRDS(Downs, file = "../data/AA_DOW/AA_DOW_SeuratObject.rds")

# reference_counts = fread("zcat < ../data/fetalBrain/Bhaduri2020/Bhaduri2020/exprMatrix.tsv.gz")
# genes = unname(unlist(reference_counts[,1]))
# reference_counts = rowsum(reference_counts[,-1], genes)
# reference_meta <- read.table("../data/fetalBrain/Bhaduri2020/Bhaduri2020/meta.tsv", header=T, sep="\t", as.is=T, row.names=1)
# genes = gsub(".+[|]", "", genes)
# Reference <- CreateSeuratObject(counts = reference_counts, project = "fetalBrain_Reference", meta.data = reference_meta)
# saveRDS(Reference, file = "../data/fetalBrain/reference_SeuratObject.rds")
Reference = readRDS('../data/fetalBrain/reference_SeuratObject.rds')
Downs = readRDS("../data/fetalBrain/Bhaduri2020/Bhaduri2020/AA_DOW_SeuratObject.rds")
```

The QC plots using number of detected genes, number of counts and percent of counts coming from mitochondrial genes (as a proxy for stress), show a couple of outlier cells, which I remove:
  
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
Downs.list <- SplitObject(Downs, split.by = 'Donor')
i = 1
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
i = 2
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
i = 3
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
i = 4
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
```

The following normalizes, scales and select 2000 particularly variable genes. 

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
Downs.list[[i]] <- NormalizeData(Downs.list[[i]], normalization.method = "LogNormalize", scale.factor = 10000)
Downs.list[[i]] <- FindVariableFeatures(Downs.list[[i]], selection.method = "vst", nfeatures = 2000)
top25 <- head(VariableFeatures(Downs.list[[i]]), 25)
plot1 <- VariableFeaturePlot(Downs.list[[i]])
plot2 <- LabelPoints(plot = plot1, points = top25, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
all.genes <- rownames(Downs.list[[i]])
Downs.list[[i]] <- ScaleData(Downs.list[[i]], features = all.genes)
}
Reference = NormalizeData(Reference, normalization.method = "LogNormalize", scale.factor = 10000)
Reference <- FindVariableFeatures(Reference, selection.method = "vst", nfeatures = 2000)
```

Integration:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}

predictions_Type = list()
predictions_Subtype = list()
for (i in 1:4){
  downs.anchors <- FindTransferAnchors(reference = Reference, query = Downs.list[[i]], dims = 1:30, verbose = TRUE)
  predictions_Type[[i]] <- TransferData(anchorset = downs.anchors, refdata = Reference$Type, dims = 1:30)
  predictions_Subtype[[i]] <- TransferData(anchorset = downs.anchors, refdata = Reference$Subtype, dims = 1:30)
}

saveRDS(prediction_Type, file = '/home/jovyan/data/AA_DOW/predictionsType.rds')
saveRDS(prediction_Type, file = '/home/jovyan/data/AA_DOW/predictionsSubType.rds')

```
