### Analyze Zika infected Glioblastoma data:

---
  output:
  html_document: default
pdf_document: default
---
  ### Finding gene modules that make cells susceptible to Zika infection
  
  ---
  title: "Finding gene modules that make cells susceptible to Zika infection"
output:
  html_document: default
github_document: default
fig_width: 12
fig_height: 4 
---
  
```{r include = FALSE}
### Load data and packages
knitr::opts_chunk$set(fig.width=24, fig.height=8) 
```

Load the required R libraries:
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
library(Seurat)
library(Matrix)
library(EnsDb.Hsapiens.v75)
library(rhdf5)
library(dplyr)
require(data.table)
```

Prior to this analysis we processed the raw sequencing data with cellranger and ran an algorithm called CellBender on it to remove ambient RNA.

Here I 1.) load the data 2.) change the gene identifiers from ensemble ids to symbols 3.) construct a metadata matrix (with donor information, pcw etc.)  4.) put the data and metadata into a Seurat object. 5.) load mitochondrial genes from a database for later use 6.) Remove doublets, by loading the doublet scores I got from an algorithm (called scrublet) I ran on the data before.
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
setwd('/home/jovyan/AA_DOW/')
# sample_info = data.frame(matrix('', 8,4)) # (info from sample tracker)
# colnames(sample_info) = c('SangerSampleID', 'Donor', 'Disease', 'PcW')
# sample_info$SangerSampleID = c('DOW8726784', 'DOW8726783', 'DOW8726787', 'DOW8726788', 'DOW8726785', 'DOW8726786', 'DOW8726789', 'DOW8726790')
# sample_info$Donor = c('Hob1', 'Hob1', 'Hob12', 'Hob12', 'Hob15', 'Hob15', 'Hob16', 'Hob16')
# sample_info$Disease = c('TS21', 'TS21', 'TS21', 'TS21', 'Control', 'Control', 'Control', 'Control')
# sample_info$PcW = c('13pcw', '13pcw', '14pcw', '14pcw', '13pcw', '13pcw', '13pcw', '13pcw')
# scrublet_cutoff = c(0.33, 0.25, 0.4, 0.4, 0.4, 0.45, 0.4, 0.3) # From inspection of the scrublet plots (not included). This is quite a stringent doublet cutoff.
# i = 1
# print(i)
# sampleID = sample_info$SangerSampleID[i]
# scrublet_score = read.delim(paste('/home/jovyan/AA_DOW/scrublet/', sampleID ,'_doublets_scores.txt', sep = ''))
# dow_counts = as.matrix(Read10X_h5(filename = paste('../data/AA_DOW/AA_DOW/cellranger302_count_33315_AA_', sampleID, '_GRCh38-3_0_0-premrna/output_filtered.h5', sep = ''), use.names = TRUE))
# dow_counts = dow_counts[,scrublet_score < scrublet_cutoff[i]]
# dow_metadata = data.frame(matrix('',dim(dow_counts)[2],4))
# colnames(dow_metadata) = c('SangerSampleID', 'Donor', 'Disease', 'PcW')
# dow_metadata$SangerSampleID = sample_info$SangerSampleID[i]
# dow_metadata$Donor = sample_info$Donor[i]
# dow_metadata$Disease = sample_info$Disease[i]
# dow_metadata$PcW = sample_info$PcW[i]
# geneNames = read.delim('/home/jovyan/data/AA_DOW/AA_DOW/cellranger302_count_33315_AA_DOW8726783_GRCh38-3_0_0-premrna/raw_feature_bc_matrix/features.tsv')
# for (i in 2:8){
#   print(i)
#   sampleID = sample_info$SangerSampleID[i]
#   scrublet_score = read.delim(paste('/home/jovyan/AA_DOW/scrublet/', sampleID ,'_doublets_scores.txt', sep = ''))
#   dow_counts_subset = as.matrix(Read10X_h5(filename = paste('../data/AA_DOW/AA_DOW/cellranger302_count_33315_AA_', sampleID, '_GRCh38-3_0_0-premrna/output_filtered.h5', sep = ''), use.names = TRUE))
#   dow_counts_subset = dow_counts_subset[,scrublet_score < scrublet_cutoff[i]]
#   dow_metadata_subset = data.frame(matrix('',dim(dow_counts_subset)[2],4))
#   colnames(dow_metadata_subset) = c('SangerSampleID', 'Donor', 'Disease', 'PcW')
#   dow_metadata_subset$SangerSampleID = sample_info$SangerSampleID[i]
#   dow_metadata_subset$Donor = sample_info$Donor[i]
#   dow_metadata_subset$Disease = sample_info$Disease[i]
#   dow_metadata_subset$PcW = sample_info$PcW[i]
#   dow_counts = cbind(dow_counts, dow_counts_subset)
#   dow_metadata = rbind(dow_metadata, dow_metadata_subset)
# }
# mitogenes = geneNames[,2][substring(geneNames[,2],1,3) == 'MT-']
# percent.mt = colSums(dow_counts[rownames(dow_counts) %in% mitogenes,])/colSums(dow_counts)
# Downs <- CreateSeuratObject(dow_counts, project = 'AA_DOW', min.cells = 0, min.features = 0)
# Downs$Barcode = colnames(dow_counts)
# Downs$SangerSampleID = dow_metadata$SangerSampleID
# Downs$Donor = dow_metadata$Donor
# Downs$Disease = dow_metadata$Disease
# Downs$percent.mt = unname(percent.mt)
# saveRDS(Downs, file = "../data/AA_DOW/AA_DOW_SeuratObject.rds")

reference_counts = fread("zcat < ../data/fetalBrain/Bhaduri2020/Bhaduri2020/primarymatrix_nonorm.txt.gz")
genes = unname(unlist(reference_counts[,1]))
reference_counts = rowsum(reference_counts[,-1], genes)
reference_meta <- read.table("../data/fetalBrain/Bhaduri2020/Bhaduri2020/meta.tsv", header=T, sep="\t", as.is=T, row.names=1)
genes = gsub(".+[|]", "", genes)
Reference <- CreateSeuratObject(counts = reference_counts, project = "fetalBrain_Reference", meta.data=reference_meta)
saveRDS(Reference, file = "../data/fetalBrain/Bhaduri2020/B")
Downs = readRDS("../data/AA_DOW/AA_DOW_SeuratObject.rds")


genes = unname(unlist(test[,1]))
test_new = rowsum(test[,-1], genes)

```

The QC plots using number of detected genes, number of counts and percent of counts coming from mitochondrial genes (as a proxy for stress), show a couple of outlier cells, which I remove:
  
```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
Downs.list <- SplitObject(Downs, split.by = 'Donor')
i = 1
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
i = 2
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
i = 3
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
i = 4
VlnPlot(Downs.list[[i]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
plot1 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(Downs.list[[i]], feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
CombinePlots(plots = list(plot1, plot2))
Downs.list[[i]] <- subset(Downs.list[[i]], subset = nFeature_RNA > 1000 & nFeature_RNA < 15000 & nCount_RNA < 10^5 & percent.mt < 0.05)
```

The following normalizes, scales and select 2000 particularly variable genes. 

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
Downs.list[[i]] <- NormalizeData(Downs.list[[i]], normalization.method = "LogNormalize", scale.factor = 10000)
Downs.list[[i]] <- FindVariableFeatures(Downs.list[[i]], selection.method = "vst", nfeatures = 2000)
top25 <- head(VariableFeatures(Downs.list[[i]]), 25)
plot1 <- VariableFeaturePlot(Downs.list[[i]])
plot2 <- LabelPoints(plot = plot1, points = top25, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
all.genes <- rownames(Downs.list[[i]])
Downs.list[[i]] <- ScaleData(Downs.list[[i]], features = all.genes)
}
```

These are the results of the PCA analysis:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
Downs.list[[i]] <- RunPCA(Downs.list[[i]], features = VariableFeatures(object = Downs.list[[i]]))
print(Downs.list[[i]][["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(Downs.list[[i]], dims = 1:2, reduction = "pca")
DimPlot(Downs.list[[i]], reduction = "pca")
DimHeatmap(Downs.list[[i]], dims = 1, cells = round(dim(Downs.list[[i]])[2])/2, balanced = TRUE)
DimHeatmap(Downs.list[[i]], dims = 1:15, cells = round(dim(Downs.list[[i]])[2])/2, balanced = TRUE)
}
```

Based on the JackStraw procedure I select 14 PCs for further analysis in all cases:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
# for (i in range(4)){
# Downs.list[[i]] <- JackStraw(Downs.list[[i]], num.replicate = 100)
# Downs.list[[i]] <- ScoreJackStraw(Downs.list[[i]], dims = 1:20)
# print(JackStrawPlot(Downs.list[[i]], dims = 1:20))
# print(ElbowPlot(Downs.list[[i]], ndims = 40))
# }
```

This is the clustering step:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
n_dimensions = 14
for (i in 1:4){
Downs.list[[i]] <- FindNeighbors(Downs.list[[i]], dims = 1:n_dimensions)
Downs.list[[i]] <- FindClusters(Downs.list[[i]], resolution = 0.5)
head(Idents(Downs.list[[i]]), 5)
}
```

Clusters roughly agree with visual seperation on a UMAP plot:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
Downs.list[[i]] <- RunUMAP(Downs.list[[i]], dims = 1:n_dimensions)
print(DimPlot(Downs.list[[i]], reduction = "umap"))
}

```

Visualize markers and annotate clusters:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
# Downs.markers <- FindAllMarkers(Downs, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
# saveRDS(Downs.markers, file = "../data/AA_DOW/Downs_markers.rds")
# Downs.markers = readRDS("../data/AA_DOW/Downs_markers.rds")
# top = Downs.markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)
# FeaturePlot(Downs, features = top$gene)
# top10 <- Downs %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
# DoHeatmap(Downs, features = top10$gene, cells = 1:dim(Downs)[2] + NoLegend())
# as.matrix(top10)
```

Vizualize a priori chosen markers:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
for (i in 1:4){
 print(FeaturePlot(Downs.list[[i]], features = c('CRYAB', 'HOPX', 'SOX2', 'HES1', 'PAX6', 'VIM', 'EOMES', 'NEUROD6', 'STMN2', 'SOX5', 'TBR1', 'BCL11B', 'SATB2', 'CALB2', 'SST', 'DLX2', 'DLX1', 'OLIG2', 'OLIG1', 'CLDN5', 'ITM2A', 'RGS5', 'CX3CR1', 'AIF1'))) 
}
```

Vizualize donor:

```{r, include = TRUE, message = FALSE, warning = FALSE, cache = FALSE, fig.height = 8, fig.width = 16}
DimPlot(Downs, reduction = "umap", group.by = 'Disease')
DimPlot(Downs, reduction = "umap", group.by = 'Donor')
```



